---
title: "Predicting Excercise Effectiveness from Accelerometer Data"
---

Synopsis: The purpose of this exercise is to find a model that accurately predicts how well a person exercised based on data collected from accelerometers attached to the belt, forearm, arm and dumbbell of a participant. 

Data Processing 
Data were gathered from http://groupware.les.inf.puc-rio.br/har. The data had already been split into a training and test data set, and the data from the training set were loaded into acc.training, and from the testing set into acc.testing

```{r cache=TRUE}

acc.training<-read.csv("pml-training.csv",na.strings=c("NA","#DIV/0!"))
acc.testing<-read.csv("pml-testing.csv",na.strings=c("NA","#DIV/0!"))

```
The data shows that the 'classe' variable is ordinal categorical and that there are several variables which represent aggregates, and contain NA's or #DIV/0. These data will not be useful in creating a learning model. I reduced the dataset to exclude these variables from the prediction, as well as the name and timestamp data. 

```{r}

no_na<-apply(acc.training,2,function(x) !any(is.na(x)))
acc.nona.training<-acc.training[, no_na] 
acc.nona.training<-acc.nona.training[,-c(1:8)]
dim(acc.nona.training)
```
Next I partitioned acc.nona.training into a training and test set to work on my final model.
```{r}
library(caret)
inTrain<-createDataPartition(y=acc.nona.training$classe,p=0.7,list=FALSE)
training<-acc.nona.training[inTrain,]
testing<-acc.nona.training[-inTrain,]

```

I worked with a random forest model set up with cross validation. 
```{r}
set.seed(1234)
control<-trainControl(method="cv",number=3,verbose=T,allowParallel=T)
model.rf<-train(classe~.,data=training, method="rf", trControl=trainControl, verbose=F)
fit<-predict(model.rf,testing)
confusionMatrix(fit, testing$classe)
```
The model is 99.4% accurate with a 95% CI of 99.21% to 99.61%
When using it on the test data, we get the following results 
```{r}
acc.testing<-acc.testing[,names(acc.nona.training[,-52])]
final_fit<-predict(model.rf,acc.testing)
final_fit
```
